---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# invest.data

<!-- badges: start -->

<!-- badges: end -->

`invest.data` provides tidy wrappers to collect and standardize investment data from online sources.

The package focuses on four workflows:

1. **Stock and crypto OHLCV time series** (Yahoo Finance, Tiingo, Binance)
2. **Dividend history** (Yahoo Finance)
3. **Market universes and benchmarks** (SP500, ASX, B3 + benchmark list)
4. **Fundamentals datasets** (DoltHub earnings repository)

Most data-retrieval functions return a standardized object:

- `data`: tibble/list with successful results
- `errors`: tibble with failed symbols/files and error details (`NULL` when no errors)

### How does it compare to other packages?

Several R packages already provide access to financial data:

- **`tidyquant::tq_get()`**  
  Flexible and general-purpose. Supports Yahoo, Tiingo, Alpha Vantage,
  and more.  
  *Difference*: `invest.data` builds on `tq_get()` but adds **query
  chunking**, **error logging**, and **consistent output structures**
  (`data` + `errors`).

- **`quantmod::getSymbols()`**  
  One of the oldest tools for financial data in R. Primarily designed
  for **xts/zoo** workflows.  
  *Difference*: `invest.data` returns **tibbles**, integrates with the
  tidyverse, and provides **crypto support**.

- **`riingo`**  
  Official Tiingo client. Useful for low-level Tiingo queries.  
  *Difference*: `invest.data` abstracts over multiple providers (Yahoo,
  Tiingo, Binance) with the **same interface** and adds error handling.

- **`BatchGetSymbols`**  
  Convenient batch retrieval of daily Yahoo Finance data.  
  *Difference*: `invest.data` is more general (intraday + crypto +
  dividends) and maintains the same return structure across all data
  types.

In short, `invest.data` focuses less on being a full-featured API
client, and more on being a **unified, resilient interface** for
fetching stock, crypto, and dividend data with minimal friction.

## Installation

You can install the development version from GitHub:

```r
# install.packages("remotes")
remotes::install_github("henrique-anatole/invest.data", dependencies = TRUE)
```

## Minimal quickstart

```{r quickstart, eval = FALSE}
library(invest.data)

# 1) Configure credentials (prefer keyring-backed setup)
set_tiingo_api_key()

# 2) Fetch a symbol universe
universe <- scrap_sp500_symbols()

# 3) Pick symbols and download daily prices
symbols <- universe |> dplyr::slice_head(n = 3) |> dplyr::pull(symbol)
prices <- load_stock_timeseries(
  symbols = symbols,
  interval = "1d",
  start_date = "2024-01-01",
  end_date = "2024-03-01"
)

# 4) Work with successful data + errors
prices$data
prices$errors
```

For a full walkthrough and troubleshooting guidance, see
`vignettes/getting-started.Rmd`.

## Main functions

### Time series and dividends

- `load_stock_timeseries()`
  - Loads stock OHLCV from Yahoo (`"1d"`) or Tiingo intraday (`"1m"` to `"12h"`)
  - Optional split adjustment for Tiingo data (`adjust_splits = TRUE`)
- `load_crypto_timeseries()`
  - Loads crypto OHLCV from Binance (currently the only supported source)
- `load_yahoo_dividends()`
  - Loads dividends from Yahoo Finance

### Symbol universes and benchmarks

- `scrap_sp500_symbols()`
- `scrap_asx_symbols()`
- `scrap_b3_symbols()`
- `create_benchmarks()`

### Fundamentals and utilities

- `get_fundamentals_data()`
  - Downloads multiple fundamentals tables from DoltHub
- `is_valid_date()`
  - Validates `YYYY-MM-DD` date strings

### Included datasets

- `classificacao_b3`: B3 company classification table (sector/subsector metadata)
- `fake_page`: sample HTML content for scraping-related tests

## API Key Setup

Some functions require API keys (e.g., **Tiingo** for stock data and **Binance** for crypto data).
Both are managed with helper functions:

* `set_tiingo_api_key()`
* `set_binance_api_key()`

Each function looks for your credentials in the following order:

1. Direct input via function arguments
2. Values already set in `options()`
3. Environment variables (`TIINGO_KEY`, `BIN_KEY`, `BIN_SECRET`)
4. Secure storage in the system keyring (`keyring` package)

If no valid key is found, an error is raised with setup instructions.

### Recommended method: keyring (secure storage)

This is the safest way to store credentials, since they are encrypted and managed by the OS.

```r
# Store keys securely (one-time setup)
keyring::key_set("tiingo_key")
keyring::key_set("binance_key")
keyring::key_set("binance_secret")

# Later, in your session just call:
set_tiingo_api_key()
set_binance_api_key()
```

### Alternative methods (less secure)

* **Direct Input (scripts / quick testing):**

  ```r
  set_tiingo_api_key("YOUR_TIINGO_KEY")
  set_binance_api_key("YOUR_BINANCE_KEY", "YOUR_BINANCE_SECRET")
  ```

* **Environment Variables (persistent, but stored in plain text):**
  Add to `~/.Renviron` or set in your shell:

  ```r
  Sys.setenv(TIINGO_KEY = "YOUR_TIINGO_KEY")
  Sys.setenv(BIN_KEY = "YOUR_BINANCE_KEY", BIN_SECRET = "YOUR_BINANCE_SECRET")
  ```

* **Options (only lasts for the current session):**

  ```r
  options(tiingo_key = "YOUR_TIINGO_KEY")
  options(bin_key = "YOUR_BINANCE_KEY", bin_secret = "YOUR_BINANCE_SECRET")
  ```

Use `keyring` whenever possible. Avoid hardcoding keys in scripts.

## Examples

### Stock OHLCV (daily via Yahoo)

Load daily prices from Yahoo Finance:

```{r example-stock, eval = FALSE}
library(invest.data)

stock_test <- load_stock_timeseries(
  symbols    = c("AAPL", "MSFT", "INVALID"),
  interval   = "1d",
  start_date = "2021-10-01",
  end_date   = "2022-02-01"
)

head(stock_test$data)
head(stock_test$errors)
```

### Stock OHLCV (intraday via Tiingo)

Load hourly intraday data (Tiingo):

```{r example-intraday, eval = FALSE}
stock_test_hour <- load_stock_timeseries(
  symbols    = c("AAPL", "MSFT", "INVALID"),
  interval   = "1h",
  start_date = "2021-10-01",
  end_date   = "2021-10-02"
)

head(stock_test_hour$data)
head(stock_test_hour$errors)
```

### Crypto OHLCV (Binance)

Load crypto OHLCV from Binance:

```{r example-crypto, eval = FALSE}
crypto_test <- load_crypto_timeseries(
  pair       = c("ETHUSDT", "BTCUSDT", "INVALID"),
  interval   = "1m",
  start_date = "2021-01-01",
  end_date   = "2021-01-02"
)

head(crypto_test$data)
head(crypto_test$errors)
```

### Dividends (Yahoo)

Load dividends from Yahoo Finance:

```{r example-dividends, eval = FALSE}
dividends <- load_yahoo_dividends(
  symbols    = c("AAPL", "MSFT", "INVALID"),
  start_date = "2020-01-01",
  end_date   = "2022-01-01"
)

head(dividends$data)
head(dividends$errors)

```

### Market universes and benchmarks

```{r example-symbols, eval = FALSE}
sp500 <- scrap_sp500_symbols()
benchmarks <- create_benchmarks()

sp500 |> dplyr::count(sector, sort = TRUE)
benchmarks |> dplyr::slice_head(n = 5)
```

### Fundamentals tables

```{r example-fundamentals, eval = FALSE}
fundamentals <- get_fundamentals_data(
  files_to_get = c("income_statement", "cash_flow_statement")
)

names(fundamentals$data)
fundamentals$errors
```

## Operational notes

- `load_*` and `get_fundamentals_data()` return a list with `data` and `errors`.
- Symbol scraping functions (`scrap_*`) return a tibble on success and may return `NULL` with a warning on failure.
- `scrap_asx_symbols()` and `scrap_b3_symbols()` require Selenium + Firefox.
- `load_stock_timeseries()` uses Tiingo for intraday intervals and Yahoo for daily.
- `load_crypto_timeseries()` currently supports only Binance (`source = "binance"`).
- Date inputs are expected in `YYYY-MM-DD`; use `is_valid_date()` for validation.

## Known limitations

- Upstream providers can enforce rate limits and may throttle or block repeated requests.
- Web scraping sources (ASX/B3 pages) can break when site layout or download flows change.
- Intraday data coverage can vary by provider/symbol and may have gaps.
- Source outages or temporary network issues can produce partial results; check `errors`.
- `RSelenium` workflows depend on local browser/driver compatibility.

## RSelenium setup (ASX/B3)

`scrap_asx_symbols()` and `scrap_b3_symbols()` depend on browser automation.

Prerequisites:

- Firefox installed on the machine
- Java runtime available (required by Selenium server)
- `RSelenium` package installed

If Selenium setup fails, start by checking browser/driver compatibility and whether the chosen port is already in use.

For production pipelines, prefer one of these approaches:

- Run scraping in an isolated process/container
- Use fixed environment setup (same Firefox + geckodriver versions)
- Cache scraped outputs and refresh on a schedule

## Reproducibility notes

- Record your R and package versions (`sessionInfo()`) in analysis outputs.
- Pin dependency versions for production jobs (for example with `renv`).
- For Selenium workflows, keep Firefox and geckodriver versions fixed in CI/containers.
- Rebuild universes and fundamentals on a defined cadence to avoid drift.
- This README was last updated on 2026-06-03.

Extended documentation is available in `vignettes/getting-started.Rmd`.

## Data sources

| Function | Primary source(s) |
|---|---|
| `load_stock_timeseries()` | Yahoo Finance (`tidyquant::tq_get`), Tiingo (`tidyquant`/`riingo`) |
| `load_crypto_timeseries()` | Binance (`binancer`) |
| `load_yahoo_dividends()` | Yahoo Finance (`tidyquant::tq_get`) |
| `scrap_sp500_symbols()` | tidyquant index + Wikipedia + StockAnalysis |
| `scrap_asx_symbols()` | MarketIndex |
| `scrap_b3_symbols()` | B3 IBOV page + internal/external classification |
| `get_fundamentals_data()` | DoltHub `earnings` repository |
