---
title: "Getting Started with invest.data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with invest.data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

`invest.data` provides wrappers for:

- stock OHLCV (`load_stock_timeseries()`)
- crypto OHLCV (`load_crypto_timeseries()`)
- dividends (`load_yahoo_dividends()`)
- symbol universes (`scrap_sp500_symbols()`, `scrap_asx_symbols()`, `scrap_b3_symbols()`)
- fundamentals datasets (`get_fundamentals_data()`)

Most retrieval functions return a list with:

- `data`: successful results
- `errors`: failed symbols/files and messages

## Installation

```{r}
# install.packages("remotes")
remotes::install_github("henrique-anatole/invest.data", dependencies = TRUE)
library(invest.data)
```

## API credentials

Use keyring-backed setup when possible:

```{r}
keyring::key_set("tiingo_key")
keyring::key_set("binance_key")
keyring::key_set("binance_secret")

set_tiingo_api_key()
set_binance_api_key()
```

## End-to-end workflow

```{r}
# 1) Fetch universe
sp500 <- scrap_sp500_symbols()

# 2) Select symbols
symbols <- sp500 |> dplyr::slice_head(n = 5) |> dplyr::pull(symbol)

# 3) Download prices
prices <- load_stock_timeseries(
  symbols = symbols,
  interval = "1d",
  start_date = "2024-01-01",
  end_date = "2024-04-01"
)

# 4) Inspect successful and failed requests
prices$data
prices$errors

# 5) Add benchmarks
benchmarks <- create_benchmarks()
benchmarks |> dplyr::slice_head(n = 5)
```

## RSelenium notes (ASX/B3)

`scrap_asx_symbols()` and `scrap_b3_symbols()` require:

- Firefox installed
- Java available
- `RSelenium` configured

```{r}
install.packages("RSelenium")

# connectivity check only (optional)
driver <- RSelenium::rsDriver(browser = "firefox", port = 4445L)
remDr <- driver$client
remDr$navigate("https://example.org")
remDr$close()
driver$server$stop()
```

For automated environments, pin browser/driver versions and run scraping in an isolated container or worker.

## Known limitations

- Data providers can enforce request limits and throttling.
- Web-scraped sources can change HTML/download patterns without notice.
- Intraday coverage depends on symbol/provider and may be incomplete.
- Temporary source/network failures can produce partial outputs.

## Reproducibility

```{r}
sessionInfo()
```

Recommended practices:

- Use `renv` (or equivalent) to lock package versions.
- For Selenium workflows, pin Firefox/geckodriver versions.
- Cache extracted universes/fundamentals and refresh on a fixed schedule.
